Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support).

### 2017-02-06 csf-2017/paper.tex
changed ordering, added a comment
small intro update

\section{Introduction}
\label{sec:introduction}
\todo[inline]{See the latex comments for the next paragraphs!}
% Possible high-level intros:
% 1) Security guarantees vs obstacles / arms race
Security on processors is today subject to an arms race between
processor manufacturers and hackers. On the hacker side, a constant
effort is put into exploiting current processor designs and
circumventing the security measures set in place to prevent known
exploits. On the processor manufacturer side, new security measures
are put in place to prevent new exploits. The arms race would come to
a grinding halt if the security meassures provided more than obstacles
and actually provided low-level security guarantees. One kind of
low-level machines that provide low-level security is a capability machine.

% 2) Memory access-control

% 3) Fully abstract compilation
High-level languages provide guarantees such as encapsulation of local
state and well-bracketedness of function calls. These guarantees are
often relied upon when reasoning about the correctness and security of
a program. High-level programs are compiled to machine code, so in
order to rely on the correctness and security guarantees provided by
the high-level language, it needs to be proven that the translation
preserves these properties. Translated programs often interact with
machine code that is not translated from the same high-level language,
so in order to preserve the guarantees from the high-level languages,
the low-level machine needs to provide some kind of security
guarantees. In other words, a target of secure compilation needs to
provide some security guarantees. Modern processors provide many
security meassures, but very little in terms of guarantees. A machine
that provides some means of security guarantees is a capability machine.

@@

% Capability machines offer fine-grained memory protection
A capability machine is a low-level machine that offers fine-grained
memory protection.
% Capability machines and unforgeable tokens for memory access. 
This is done by introducing unforgeable tokens to
the machine. These tokens are called capabillities and grant a kind of
authority and a range of authority. We will refer to the capabilities
that grant a combination of read, write, or execute permission as memory
capabilities. Capabilities with the necessary permissions must be
presented every time an instruction that manipulates the memory is
executed.

@@

Capabilities are irrevocable, so when we pass a capability has to
an untrusted program, then we have to assume that the program keeps it
around indefenitely. This means that we cannot reuse the piece of
memory that the capability governs essentially creating a memory
leak. The CHERI processor has a special kind of capabilities called
\emph{local capabilities}\todo{reference?} that introduces a simple
kind of temporal information control which allows for a simple kind of
capability revocation. This is achieved by adding a tag to every
capability that marks it as either local or global. Local capabilities
can only be written through a capability with a new permission called
\emph{permit write local}.
% More about local capabilities.

>>>>

\section{Introduction}
\label{sec:introduction}
\todo[inline]{See the latex comments for the next paragraphs!}
% Possible high-level intros:
% 0) Arms race compilers
%
% 1) Security guarantees vs obstacles / arms race
\begin{comment}
  Security on processors is today subject to an arms race between
  processor manufacturers and hackers. On the hacker side, a constant
  effort is put into exploiting current processor designs and
  circumventing the security measures set in place to prevent known
  exploits. On the processor manufacturer side, new security measures
  are put in place to prevent new exploits. The arms race would come
  to a grinding halt if the security meassures provided more than
  obstacles and actually provided low-level security guarantees. One
  kind of low-level machines that provide low-level security is a
  capability machine.
\end{comment}
% 2) Fully abstract compilation
High-level languages provide guarantees such as encapsulation of local
state and well-bracketedness of function calls. These guarantees are
often relied upon when reasoning about the correctness and security of
a program. High-level programs are compiled to machine code, so in
order to rely on the correctness and security guarantees provided by
the high-level language, it needs to be proven that the translation
preserves these properties. Translated programs often interact with
machine code that is not translated from the same high-level language,
so in order to preserve the guarantees from the high-level languages,
the low-level machine needs to provide some kind of security
guarantees. In other words, a target of secure compilation needs to
provide some security guarantees. Modern processors provide many
security meassures, but very little in terms of guarantees. A machine
that provides some means of security guarantees is a capability machine.

% 3) Memory access-control
% I don't like this angle and can't find a good way to sell it.

% 4) History
% Proposals throughout history.
% Lack of formal account of properties-guarantees.

@@

% Capability machines offer fine-grained memory protection
% Too technical
A capability machine is a low-level machine that offers fine-grained
memory protection.
% Capability machines and unforgeable tokens for memory access. 
This is done by introducing unforgeable tokens to
the machine. These tokens are called capabillities and grant a kind of
authority and a range of authority. 
% Dynamic checks encapsulation, formal model

 We will refer to the capabilities
that grant a combination of read, write, or execute permission as memory
capabilities. 
 Capabilities with the necessary permissions must be
presented every time an instruction that manipulates the memory is
executed.

@@

Capabilities are irrevocable, so when we pass a capability has to
an untrusted program, then we have to assume that the program keeps it
around indefenitely. This means that we cannot reuse the piece of
memory that the capability governs essentially creating a memory
leak. The CHERI processor has a special kind of capabilities called
\emph{local capabilities}\todo{reference?}\ that introduces a simple
kind of temporal information control which allows for a simple kind of
capability revocation. This is achieved by adding a tag to every
capability that marks it as either local or global. Local capabilities
can only be written through a capability with a new permission called
\emph{permit write local}.
% More about local capabilities.
